### 2025-07-18 base-rates.qmd
20250717 - chapter overviews
20250718 - cross-references

@@

Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
The calculation of posttest probability is described in INSERT.
A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
The updated (posttest) probability is where the connecting line crosses the third, right line.

@@

The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.

>>>>

This chapter provides an overview of base rates and the important roles they play in prediction.

@@

Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
The calculation of posttest (posterior) probability is described in @sec-posttestProbability.
A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
The updated (posttest) probability is where the connecting line crosses the third, right line.

@@

The blue line indicates the [posterior probability](#sec-posttestProbability) of the condition given a positive test.
The pink line indicates the [posterior probability](#sec-posttestProbability) of the condition given a negative test.
